---
title: "SynopFigures_2022-04-20"
author: "Kriddie"
date: '2024-02-10'
output: html_document
---

#About
this R markdown contains scripts to create figure in manuscript: 
"Carbon Evasion Dynamics in a Tropical, High-Elevation, Peatland Catchment are Mediated by A Threshold in Watershed Area"

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(readr)
library(tidyr) # for `gather`
library(ggplot2)
library(cowplot)
library(tidyverse)
library(rstatix)
library(ggpubr)
library(plotly)
library(gridExtra)
library(grid)
library(ggmap)
library(dplyr)
library(scales)
library(olsrr)
library(scales)

```


*Objective:*
The objective of this study is to identify high resolution spatial patterns in pCO2, CO2 flux and geomorphology within a Paramo river network.

Hypothesis 1: Off-gassing from upstream reaches and decreased connection with groundwater will result in decreasing pCO2 within a river reach from upstream to downstream and within a river network from smaller to larger streams. 

Hypothesis: 2: Higher turbulence in larger streams will yield flux rates than smaller streams

#read in data


```{r set colors for sites}

#set colors for sites

#new api
register_stadiamaps("e723ef39-9560-41be-969c-36870f7b7f84", write = TRUE)

#extract hex color codes for a plot with three elements in ggplot2 
hex1 <- hue_pal()(1)
hex3 <- hue_pal()(3)
hex6 <- hue_pal()(6)
hex8 <- hue_pal()(8)

antens <- hex1[1] #antenas
gavi <- hex3[3] #gavilan
gavi_in <- hex6[4] #gavilan inlet
gavi_out <- hex6[5] #gavilan outlet
gavitrib1 <- hex8[2] #gavilan trib1
gavitrib2 <- hex8[3] #gavilan trib2
gavitrib3 <- hex8[4] #gavilan trib3
gavitribs <- hex3[2] #gavi tribs
colmillo <- hex8[8] #colmillo
```

#fig 1
Site map generated using arcpro, not included here

#fig 2
Show general trend in co2
river network and reach

figure panels plotted separately 

```{r figure 2, echo=FALSE}
df<- read_csv(here::here("Whitmore_et_al_2024/Data.csv"))

air_ppm <- 416.45

MyTheme <-  theme_bw() + 
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16)) +
  theme(legend.text=element_text(size=16))+
  theme(aspect.ratio=1)

#figure 2 a
Fig2a <- ggplot(df %>%filter(Wetland_3=="Gavilan River Network"), aes(x = CatchmentSize_ha, y = pCO2_ppm)) +  geom_point(size=3) +
  ylab(expression(paste(italic('p'),"CO"[2] ," [ppm]")))+xlab("Catchment Size [Hectares]") +
  scale_y_log10() + geom_smooth(method='lm') + MyTheme + 
  annotate("text", x = 95, y = 10000, label  = expression("p-value < .0001\n", r^2 ~ "=" ~ .314), size = 4)


#figure 2 b
df <- df[order(df$pCO2_ppm),]

Fig2b <- qmplot(lon_fit, lat_fit, data = df %>%filter(Wetland_3=="Gavilan River Network")%>%drop_na(pCO2_ppm),size= 6, maptype = "stamen_terrain", color = log10(pCO2_ppm)) +
  scale_color_gradient(low="blue", high="red", breaks=c(log10(300),log10(1000), log10(3000), log10(10000)), labels=c("300","1,000","3,000","10,000"))+
  guides(size = "none") +
  theme(legend.text=element_text(size=14)) + 
  labs(color=expression(paste(italic('p'),"CO"[2] ," [ppm]"))) +
  theme(legend.title=element_text(size=16), legend.text=element_text(size=14)) 


### #figure 2 c.

#annotate text 
dat_text <- data.frame(
  label = c("p-value = 0.011", "p-value < .0001", 
            "p-value = 0.027","p-value = 0.11"),
  Wetland_4   = c("ANTE", "Gavilan Outlet", "GAVItrib1", "GAVItrib2"),
  Wetland = c("ANTE","Gavi-mainstem","GAVItrib1","GAVItrib2"))

#subtract 523m from outlet
df[df$Wetland_4 == 'Gavilan Outlet',]$dist <- df[df$Wetland_4 == 'Gavilan Outlet',]$dist - 523
Wetland_4_names <- c("ANTE" = "Antenas",
                    "Gavilan Outlet" = "Gavil\u00e1n Outlet",
                    "GAVItrib1" = "Gavil\u00e1n Tributary 1",
                    "GAVItrib2" = "Gavil\u00e1n Tributary 2")

#plot stream reaches where significant relationship found
Fig2c_significant <- ggplot(df %>%drop_na(Wetland)%>%filter(Wetland_4!="Gavilan Inlet")%>%filter(Wetland!="COLM")%>%filter(Wetland!="GAVItrib3"), aes(x = dist, y = pCO2_ppm)) +
  geom_point(size=3) +  
  scale_y_log10() +
  facet_grid(Wetland_4~.,labeller = as_labeller(Wetland_4_names)) + geom_smooth(method='lm')+
  ylab(expression(paste(italic('p'),"CO"[2] ," [ppm]"))) + xlab("Distance downstream [m]") +
  theme_bw() + theme(text = element_text(size = 14)) +
  geom_text(data = dat_text, mapping = aes(x = 400, y = 9000, label = label))


#plot stream reaches where no significant relationship was found

Wetland_4_names2 <- c("Gavilan Inlet" = "Gavil\u00e1n Intlet",
                    "COLM" = "Colmillo")

#annotate text 
dat_text2 <- data.frame(
  label = c("p-value = 0.35", "p-value = 0.82"),
  Wetland_4   = c("Gavilan Inlet", "COLM"),
  Wetland = c("Gavi-mainstem","COLM"))

Fig2c_notsignificant <- ggplot(df%>%drop_na(Wetland)%>%filter(Wetland_4=="Gavilan Inlet"|Wetland_4=="COLM"), aes(x = dist, y = pCO2_ppm)) + geom_point(size=3) + 
  scale_y_log10() + expand_limits(y = c(200, 4000)) +
  facet_grid(Wetland_4~.,labeller = as_labeller(Wetland_4_names2)) +
  ylab(expression(paste(italic('p'),"CO"[2] ," [ppm]"))) + xlab("Distance downstream [m]") +
  theme_bw() +
  theme(text = element_text(size = 14)) +
  geom_text(data    = dat_text2, mapping = aes(x = 900, y = 3000, label = label))

```


#figure 3

Box plot comparisons between stream reaches


```{r fig 4 boxplots, echo=FALSE}
df<- read_csv(here::here("Whitmore_et_al_2024/Data.csv"))

air_ppm <- 416.45
###box and wisker####

MyTheme <-  theme_bw() + 
  theme(axis.text.x = element_blank(),
    axis.ticks.x = element_blank(),
    axis.title.x = element_blank()) +
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=14)) +
  theme(legend.text=element_text(size=16))


Fig3a <- ggplot(df, aes(x = Wetland_5, y = pCO2_ppm, fill = Wetland_5)) + geom_boxplot() + geom_jitter(shape=18,position=position_jitter(0.1)) + 
  stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, color = "black",fill="white")+
  geom_hline(yintercept= air_ppm,color="red",linetype="twodash", linewidth=1) +
  scale_fill_manual(name=NULL,
    values = c(antens,colmillo, gavi_in,gavi_out,gavitribs),
   labels=c("Antenas", "Colmillo","Gavil\u00e1n Inlet", "Gavil\u00e1n Outlet","Gavil\u00e1n Tributaries"))+
  ylab(expression(paste(italic('p'),"CO"[2]," [ppm]"))) +
  scale_y_log10() + MyTheme +
  annotate("text", x="ANTE", y= 18000,label="A", size = 4) +
  annotate("text", x="COLM", y= 18000,label="A", size = 4) +
  annotate("text", x="Gavilan Inlet", y= 18000,label="A", size = 4) +
  annotate("text", x="Gavilan Outlet", y= 18000,label="AB", size = 4) +
  annotate("text", x="GavilanTribs", y= 18000,label="B", size = 4)


Fig3b <- ggplot(df%>%drop_na(F_CO2_umol_m2_s), aes(x = Wetland_5, y = log1p(F_CO2_umol_m2_s), fill = Wetland_5)) + 
  geom_boxplot() + geom_jitter(shape=18,position=position_jitter(0.1)) +
   stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, color = "black",fill="white")+
#  scale_y_log10() + 
  MyTheme +
  scale_fill_manual(name=NULL,
    values = c(antens,colmillo, gavi_in,gavi_out,gavitribs),
    labels=c("Antenas", "Colmillo","Gavil\u00e1n Inlet", "Gavil\u00e1n Outlet","Gavil\u00e1n Tributaries")) +
    ylab(expression(paste(" Evasion [",mu,"mol ", m^-2, s^-1,"]"))) +
    annotate("text", x="ANTE", y= log1p(5),label="A", size = 4) +
    annotate("text", x="COLM", y= log1p(5),label="B", size = 4) +
  annotate("text", x="Gavilan Inlet", y= log1p(5),label="AB", size = 4) +
  annotate("text", x="Gavilan Outlet", y= log1p(5),label="AB", size = 4) +
    annotate("text", x="GavilanTribs", y= log1p(5),label="B", size = 4) 


Fig3c <- ggplot(df%>%drop_na(K600), aes(x = Wetland_5, y = K600, fill = Wetland_5)) + 
  geom_boxplot() + geom_jitter(shape=18,position=position_jitter(0.1)) +
   stat_summary(fun = "mean", geom = "point", shape = 23, size = 3, color = "black",fill="white") +
  scale_y_log10() +
  scale_fill_manual(name=NULL,
    values = c(antens,colmillo, gavi_in,gavi_out,gavitribs),
    labels=c("Antenas", "Colmillo","Gavil\u00e1n Inlet", "Gavil\u00e1n Outlet","Gavil\u00e1n Tributaries"))+
  ylab(expression(paste(italic('k')[600]," [m ", d^-1,"]"))) +
  MyTheme + 
  annotate("text", x="ANTE", y= 155,label="A", size = 4) +
  annotate("text", x="COLM", y= 155,label="B", size = 4) +
  annotate("text", x="Gavilan Inlet", y= 155,label="AB", size = 4) +
  annotate("text", x="Gavilan Outlet", y= 155,label="B", size = 4) +
  annotate("text", x="GavilanTribs", y= 155,label="A", size = 4)

require(grid)  



Fig3 <- ggarrange(Fig3a , Fig3b,  Fig3c,
                    ncol = 1, nrow = 3,
                    common.legend = TRUE, legend = "right",
                    align = "hv", 
                    font.label = list(size = 20, color = "black", face = "bold", family = NULL, position = "top"))

Fig3



```

#Figure 4
figure 4 show the relationship between slope and pCO2

```{r fig 4 boxplots, echo=FALSE}


df_slope <- read.csv(here::here("Whitmore_et_al_2024/Slope_Data.csv"))

theme_set(theme_bw(base_size = 16))

fig4 <- ggplot(df_slope  %>%filter(CatchmentSize_ha > 12.5), 
       aes(x=slope_m.m,y=pCO2_ppm, color=CatchmentSize_ha)) + 
  geom_point(size=2) +
  geom_smooth(method = "lm",color="black") +
  ylab(expression(paste(italic('p'),"CO"[2] ," [ppm]"))) + xlab("Slope [m/m]") +
  scale_y_continuous(trans = "log10", breaks=c(300,1000,3000,10000), labels=c("300","1,000","3,000","10,000")) +
  scale_colour_gradient( name = "Catchment Size [ha]",
  low = "yellow3", high = "green4",
  space = "Lab", na.value = "grey50",
  guide = "colourbar",aesthetics = "colour")


```

#figure 5
figure 5 shows the relationship between catchment size and k600

```{r fig 5 , echo=TRUE}

df_slope <- read.csv(here::here("Whitmore_et_al_2024/Slope_Data.csv"))

fig5 <- ggplot(df_slope, 
            aes(x=CatchmentSize_ha,y=K600)) +
  scale_y_log10() + scale_x_log10() +
  geom_point(size=2) +
  geom_smooth(method = "lm",color="black") +
  ylab(expression(paste(italic('k'),"600 [m d" ^1,"]"))) + xlab("Catchment Size [ha]")
fig5

```
#figure 6

figure 6 show drivers of evasion in small large
for < 12.5 hectars, watershed size is not significant, co2 is significant
r2 = 0.5247, pvalue =2.43e-07
for > 20 ha, co2 is not significant, but watershed size is. pvalue = 0.00395 r2 = 0.1057

```{r fig 6, echo=FALSE}
df<- read_csv(here::here("Whitmore_et_al_2024/Data.csv"))

air_ppm <- 416.45

MyTheme <-  theme_bw() + 
  theme(axis.text=element_text(size=14),
        axis.title=element_text(size=16)) +
  theme(legend.title = element_text(size=18),
    legend.text=element_text(size=16))

hex <- hue_pal()


Fig6a <- ggplot(df%>%filter(CatchmentSize_ha<12.5), aes(x=pCO2_ppm, y=F_CO2_umol_m2_s)) +
  geom_point(aes(fill=Wetland),  color="black",pch=21, size=3) + 
  geom_smooth(method='lm') +
  geom_vline(xintercept= air_ppm,color="red",linetype="twodash", linewidth=1) +
  scale_y_log10() + scale_x_continuous(trans = "log10", breaks=c(300,1000,3000,10000), labels=c("300","1,000","3,000","10,000")) +
  xlab("CO2 ppm") +  ylab(expression(paste("CO"[2]," Flux [",mu,"mol ", m^-2, s^-1,"]"))) +
  labs(linetype="LEGEND", color="LEGEND") +
  scale_fill_manual(name="Stream Reach",labels=c("Antenas", "Gavil\u00e1n Tributary 1","Gavil\u00e1n Tributary 3"), values = c(antens, gavitrib1, gavitrib3)) +
  annotate("text", x = 1000, y = 2.1, label  = expression("p-value < .0001\n", r^2 ~ "=" ~ .48), size = 4) +
   theme_bw()  + MyTheme 


Fig6b <- ggplot(df %>% filter(CatchmentSize_ha>20), aes(x=CatchmentSize_ha, y=F_CO2_umol_m2_s)) +
  geom_point(aes(fill=Wetland), color="black",pch=21, size=3) + 
  scale_y_log10() + scale_x_log10() +
  geom_smooth(method='lm') +
  xlab("Catchment Size [ha]") +  ylab(expression(paste("CO"[2]," Flux [",mu,"mol ", m^-2, s^-1,"]"))) + 
  labs(linetype="LEGEND", color="LEGEND") +
  scale_fill_manual(name="Stream Reach",
    labels=c( "Colmillo","Gavil\u00e1n","Gavil\u00e1n Tributary 2"),
    values = c(colmillo, gavi,gavitrib2))+
  annotate("text", x = 60, y = 2.3, label  = expression("p-value = .0004\n", r^2 ~ "=" ~ .17), size = 4) +
   theme_bw()  + MyTheme



Fig6 <- ggarrange(Fig6a, Fig6b,  
                    ncol = 1, nrow = 2,
                    legend = "right",
                    align = "hv", 
                    font.label = list(size = 25, color = "black", face = "bold", family = NULL, position = "top"))



```


#Fig 7 
This figure show cumulative co2 evasion


```{r Figure 7 fig, echo=FALSE}

fig7_data <- read.csv(here::here("Whitmore_et_al_2024/fig7_dataframe.csv"))

MyTheme <-  theme_bw() + 
  theme(axis.text=element_text(size=18),
        axis.title=element_text(size=24)) +
  theme(legend.title = element_text(size=24),
    legend.text=element_text(size=18))+
  theme(aspect.ratio=1)

Fig7a <- ggplot(fig7_data%>%filter(Wetland_4 != "Gavilan Outlet")%>%filter(Wetland_4 != "Gavilan Inlet"), aes(x=dist, y=Totalflux_umolpers, group=Wetland_4)) +
  geom_line(aes(color=Wetland), size=3) +
  geom_point(aes(shape = shape,fill=Wetland), size=2) +
  ylab(expression(paste(""))) + xlab("") +
  scale_color_manual(name="Stream Reach",labels=c("Antenas", "Colmillo","Gavilan-mainstem","Tributary 1","Tributary 2"), values=c(antens, colmillo, gavi,gavitrib1, gavitrib2)) +
  scale_fill_manual(name="Stream Reach",labels=c("Antenas", "Colmillo","Gavilan-mainstem","Tributary 1","Tributary 2"), values=c(antens, colmillo, gavi,gavitrib1, gavitrib2)) +
  scale_linetype_discrete(name="Stream Reach",labels=c("Antenas", "Colmillo","Gavilan-mainstem","Tributary 1","Tributary 2")) +
  scale_shape_manual(name="Sample Point", labels=c("",""), values = 21) +
  guides(fill=FALSE,color=FALSE,shape=FALSE) +
  MyTheme

#zoom in
Fig7b <- ggplot(fig7_data%>%filter(Wetland_4 != "Gavi-mainstem")
       , aes(x=dist, y=Totalflux_umolpers, group=Wetland_4)) +
  geom_line(aes(color=Wetland_4), size=3) +
  geom_point(aes(shape = shape,fill=Wetland_4), size=2) +
  ylab(expression(paste("Cumulative CO"[2]," Flux [",mu,"mol ", s^-1,"]"))) + 
  xlab("Distance [m]") +
  scale_color_manual(name="Stream Reach",
    labels=c("Antenas", "Colmillo","Gavilan Inlet","Gavilan Outlet","Tributary 1","Tributary 2"),
    values=c(antens, colmillo, gavi_in,gavi_out,gavitrib1, gavitrib2)) +
  scale_linetype_discrete(
    name="Stream Reach",
    labels=c("Antenas", "Colmillo","Gavilan Inlet","Gavilan Outlet","Tributary 1","Tributary 2")) +
  scale_shape_manual(name="Sample Point", 
    labels=c("",""), values = 21) +
  scale_fill_manual(
values=c(antens, colmillo, gavi_in,gavi_out,gavitrib1, gavitrib2)) +
    guides(fill=FALSE) +
  xlim(-10, 350) +
  ylim(-20, 300) +
  MyTheme +
  theme(legend.position = "none") 


# Extract the legend. Returns a gtable
leg <- get_legend(Fig7a)
leg2 <- get_legend(Fig7b)

# Convert to a ggplot and print
as_ggplot(leg)
as_ggplot(leg2)


```


